Galois Characterization of Function Fields of \\ Transcendence Degree $1$ over Number Fields

### 2017-06-04 Master Thesis/synopsis.tex
Clarify uniqueness of phi in proof of Pop's Theorem

\begin{definition}\label{4.4}
For a finite subset $S\subset X_0$ set:
\[ K^S=\{x\in K^\times \mid v_P(x)=0\text{ for all }P\in X_0 -  S \} \]
$K'^{S'}$ is defined similarly. For finite subsets $S$ of $X_0$, the corresponding subsets of $X_0'$ will be denoted by $S'$, i.e. $S'=\{\sigma P\mid P\in S \} $.
\end{definition}

@@

\begin{proof}
By Lemma \ref{5.4}, the following maps are injective for all admissible subsets $S\subset X_0$:
\[ r_S:K^S \to\prod_{P\in X_0 -  S} k(P)^\times,\ x\mapsto (r_P(x))_P \]
Now it's evident that $\varphi$ is unique. For existence, let $(S,T)$ be an admissible pair. First, we show that there is a unique homomorphism $\varphi^{S,T}$, making the following diagram commutative:
\[ (\star) \qquad \begin{tikzcd}
K'^{S'} \rar["\varphi^{S,T}", dashed] \dar["r_{S',T'}"'] & K^S \dar["r_{S,T}"]\\
\displaystyle \prod_{P\in T}k(\sigma P)^\times \rar["\prod\varphi_P"'] & \displaystyle \prod_{P\in T}k(P)^\times
\end{tikzcd} \phantom{\qquad (\star)} \]
The vertical maps are injective by definition. Thus, it suffices to show that the following composition is trivial:
\[\varrho:K'^{S'}\longrightarrow\prod_{P\in T} k(\sigma P)^\times \longrightarrow \prod_{P\in T} k(P)^\times\longrightarrow\operatorname{coker}(r_{S,T})\] 
Consider the commutative diagram which results from tensoring $(\star)$ with $\mathbb{Z}/n\mathbb{Z}$:
\[ \begin{tikzcd}
K'^{S'}/n \rar[dashed]\dar & K^S/n\dar\\
\displaystyle \prod_{P\in T}k(\sigma P)^\times/n \rar & \displaystyle \prod_{P\in T}k(P)^\times/n
\end{tikzcd} \]
According to Lemma~\ref{4.3} and Lemma~\ref{4.5} (ii) this diagram can now be completed by $\varphi_n: K'^{S'}/n\to K^S/n$. This implies $\varrho/n=0$, i.e. $\operatorname{im}(\varrho)\subset n\operatorname{coker}(r_{S,T})$ for every $n\in\mathbb{N}$. By Lemma~\ref{5.2} (iii) and Lemma~\ref{5.4} it follows $\varrho = 0$. By symmetry, $\varphi^{S,T}$ is an isomorphism.

@@

The commutativity of the diagram follows from the commutativity of $(\star)$ with $T=X_0 -  S$. Note that with Lemma \ref{4.5} (i), we can find for all $x\in U_{P'}$ an admissible set $S\subset X_0 -  \{P\}$ such that $x\in K'^{S'}$.
\end{proof}

>>>>

\begin{definition}\label{4.4}
For a finite subset $S\subset X_0$ set:
\[ K^S=\{x\in K^\times \mid v_P(x)=0\text{ for all }P\in X_0 -  S \} = \bigcap_{P\in X_0 -  S} U_P \]
$K'^{S'}$ is defined similarly. For finite subsets $S$ of $X_0$, the corresponding subsets of $X_0'$ will be denoted by $S'$, i.e. $S'=\{\sigma P\mid P\in S \} $.
\end{definition}

@@

\begin{proof}
By Lemma \ref{5.4}, the following maps are injective for all admissible subsets $S\subset X_0$:
\[ r_S:K^S \to\prod_{P\in X_0 -  S} k(P)^\times,\ x\mapsto (r_P(x))_P \]
Using the desired commutativity of the diagram above and, with Lemma \ref{4.5} (i), the fact that for all $x\in K^\times$ there is an admissible set $S\subset X_0$ such that $x\in K^S$, it's evident that $\varphi$ is unique.

For the existence of $\varphi$, let $(S,T)$ be an admissible pair. First, we show that there is a unique homomorphism $\varphi^{S,T}$, making the following diagram commutative:
\[ (\star) \qquad \begin{tikzcd}
K'^{S'} \rar["\varphi^{S,T}", dashed] \dar["r_{S',T'}"'] & K^S \dar["r_{S,T}"]\\
\displaystyle \prod_{P\in T}k(\sigma P)^\times \rar["\prod\varphi_P"'] & \displaystyle \prod_{P\in T}k(P)^\times
\end{tikzcd} \phantom{\qquad (\star)} \]
The vertical maps are injective by definition. Thus, it suffices to show that the following composition is trivial:
\[\varrho:K'^{S'}\longrightarrow\prod_{P\in T} k(\sigma P)^\times \longrightarrow \prod_{P\in T} k(P)^\times\longrightarrow\operatorname{coker}(r_{S,T})\] 
Consider the commutative diagram which results from tensoring $(\star)$ with $\mathbb{Z}/n\mathbb{Z}$:
\[ \begin{tikzcd}
K'^{S'}/n \rar[dashed]\dar & K^S/n\dar\\
\displaystyle \prod_{P\in T}k(\sigma P)^\times/n \rar & \displaystyle \prod_{P\in T}k(P)^\times/n
\end{tikzcd} \]
According to Lemma~\ref{4.3} and Lemma~\ref{4.5} (ii) this diagram can now be completed by $\varphi_n: K'^{S'}/n\to K^S/n$. This implies $\varrho/n=0$, i.e. $\operatorname{im}(\varrho)\subset n\operatorname{coker}(r_{S,T})$ for every $n\in\mathbb{N}$. By Lemma~\ref{5.2} (iii) and Lemma~\ref{5.4} it follows $\varrho = 0$. By symmetry, $\varphi^{S,T}$ is an isomorphism.

@@

The commutativity of the desired diagram follows from the commutativity of $(\star)$ with $T=X_0 -  S$. Note that with Lemma \ref{4.5} (i), we can find for all $x\in U_{P'}$ an admissible set $S\subset X_0 -  \{P\}$ such that $x\in K'^{S'}$.
\end{proof}
